A performance comparison of Dask and Spark for data-intensive neuroimaging pipelines

### 2019-07-27 latex/paper.tex
typo
Worked on Discussion section

\subsection{Histogram: Number of blocks}
Figure~\ref{fig:histo_ms_block} shows the makespan for different block size. Again,
Spark is not compared at 30 blocks due to its \SI{2}{\giga\byte} task limitation.
Once more, Spark is significantly faster then Dask engines and Bag is slightly faster
than Delayed. Overall, the engines do not react to change in block size.

@@

\subsection{Effect of Network File System}

\subsection{Effect of data transfer}

\subsection{Effect of the Global Interpreter Lock}

\subsection{Effect of the number of partitions}

>>>>

\subsection{Histogram: Number of blocks}
Figure~\ref{fig:histo_ms_block} shows the makespan for different block size. Again,
Spark is not compared at 30 blocks due to its \SI{2}{\giga\byte} task limitation.
Once more, Spark is significantly faster than Dask engines and Bag is slightly faster
than Delayed. Overall, the engines do not react to change in block size.

@@

\subsection{Effect of Network File System}
We used an NFS as it is a common approach to store a large amount of data. However,
it limits the available bandwidth since all workers interact with one disk. This
causes data-intensive applications to take a significant amount of time for IO. As we
expect, this effect is amplified when too many tasks simultaneously try to operate
with the NFS at the same time as
Fig.~\ref{fig:inc_gantt},\ref{fig:histo_gantt},\ref{fig:histo_np_gantt},\ref{fig:bids_gantt}
shows. We could use a parallel shared file system, like Lustre \MD{cite lustre if not
done already}, to reduce the impact of operations with the disk, however, we would
quickly be limited by the network's bandwidth.

\subsection{Effect of data-intensive applications}
In our experiments, Incrementation, Histogram (NumPy) and Histogram for Spark have
low compute time while BIDS App example and Histogram for Dask have significantly
longer compute time. When adding additional workers, we observe that data-intensive
applications scale poorly while compute-intensive applications scale almost linearly
This is due to the limitation from the NFS, as mentioned above.

Data transfer accounts for the majority of the execution time for data-intensive
applications (see
Fig.~\ref{fig:inc_gantt},\ref{fig:histo_spark_gantt},\ref{fig:histo_np_gantt}). We
see that when a low number of IO tasks are overlapping, this diminishes the bandwidth
limitation from the NFS. We want to explore solutions to desynchronize IO tasks in
the aim to reduce the makespan of data-intensive applications.

\subsection{Effect of the Global Interpreter Lock}
Python's GIL prevents multi-threading in Python. Dask tries to minimize the
consequences of the GIL by using external Python extensions. We found that Dask has
issues to parallelize Python functions while Spark did not (see
Fig.~\ref{fig:histo_tt_worker}). However, Dask was easily able to parallelize
application with functions in Numpy (see
Fig.~\ref{fig:inc_tt_worker},\ref{fig:histo_np_tt_worker}) and command-line
applications (see Fig.~\ref{fig:bids_tt_worker}). This is because, for Dask, the
instances are configured to have 1 worker process with 8 threads. Since the Histogram
implementation does not release the GIL it can only use one of the thread at a given
time thus it slows down the computation significantly.

One should be careful to configure the workers properly depending on the type
of functions executed in an application.

\subsection{Effect of tasks' size}
We measured the effect of various block sizes for constant BigBrain image for
Incrementation, Histogram, and Histogram (NumPy). We observe that larger block size
results in more IO compensated by lower overhead time. On the other hand, smaller
block size results in higher overhead time compensated by reduced IO time.

The number of partitions also has an impact on the size of the tasks. We set a
partition for each block or subject. There is a trade-off for the number of
partitions. On the one hand, a low amount of partition results in larger tasks thus
in lower scheduling overhead. On the other hand, a higher number of partition
results in smaller tasks hence more tasks fit in the workers' memory insuring that
they always have something to work on.

Varying the block size is equivalent to varying the size of the partitions. We
observed no difference when varying the task size while using an NFS. However, if
one uses a parallel shared file system, we suppose it would be preferable to use a
larger block size as the IO time should be reduced substantially. It is worth noting
that the Dask developers suggest, in
\href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to
use a partition size such that workers have 2-3 times as many chunks available to
work on. More experiments would have to be performed to gain insight on an optimal
number of partitions to use for a given set of data-intensive applications.

\subsection{Other comments}
Spark performance should be reduced by the serialization of Python functions to Java,
however, our results show no significant impact of it.
